---
title: "Quarto Notebook Example"
format: html
---

```{r setup}
#| echo: false
#| include: false
```

This section will describe the **Quarto notebook** example in the repo's `{r} repo_link("examples folder", "examples/quarto/quarto_example.qmd")`.

$$
\eps\in\R
$$

$$
\y=\X\betaB, \beta
$$

# Header

The top of the file has this section of YAML delimited by 3 dashes above and below:

```
---
title: Quarto Example
author: "Jim Urick"
format:
  html:
    embed-resources: true
    monofont: Fira Code
    html-math-method: mathjax
    include-in-header:
      - text: <script type="text/javascript" src="./latex.js"></script>
---
```

You can find lots of options in the [HTML Basics](https://quarto.org/docs/output-formats/html-basics.html) documentation page. The ones I used are

- [embed-resources](https://quarto.org/docs/output-formats/html-basics.html#self-contained) renders the qmd file to a single HTML file, which includes dependencies like javascript or image files. This makes it easier to share.
- [monofont](https://quarto.org/docs/output-formats/html-themes.html#basic-options) to use Fira Code
- [html-math-methods](https://quarto.org/docs/output-formats/html-basics.html#latex-equations) specifies [MathJax](https://www.mathjax.org/) as the engine for rendering math
- [include-in-header](https://quarto.org/docs/output-formats/html-basics.html#includes) adds 


# Code Chunks and Plots

```{r}
head(mtcars)
```



```{r}
suppressPackageStartupMessages(library(tidyverse))
theme_set(theme_bw())

df <- mtcars
df |>
  ggplot(aes(mpg, disp)) +
  geom_point()
```




# TODO: image preview example

- `{r} image_preview_on_hover("image link text", "img/does_not_exist.PNG")`



## References

- $\LaTeX$ and MathJax
  - [$\LaTeX$ Cheat Sheet](https://wch.github.io/latexsheet/)
  - [The Comprehensive $\LaTeX$ Symbol List](https://us.mirrors.cicku.me/ctan/info/symbols/comprehensive/symbols-a4.pdf)
  - MathJax documentation for [defining TeX macros](https://docs.mathjax.org/en/latest/input/tex/macros.html) in javascript
- R
  - [R for Data Science](https://r4ds.hadley.nz/) is a good intro to `dplyr` and several related packages.
  - [ggplot2](https://ggplot2-book.org/)
  - [Julia Silge's YouTube](https://www.youtube.com/@JuliaSilge/videos) has lots of good ~30min videos that take a data set, do some cleaning up, and then apply machine learning models with [tidymodels](https://www.tidymodels.org/)
- Other
  - [Markdown Cheat Sheet](https://www.markdownguide.org/cheat-sheet/)
  - The Quarto website's [HTML Basics](https://quarto.org/docs/output-formats/html-basics.html) includes a bunch of links from above.
  - [Install Fira Code](https://github.com/tonsky/FiraCode/wiki/RStudio-instructions)



