---
title: "Quarto Notebook Example"
format: html
---

```{r setup}
#| echo: false
#| include: false
```

This section will describe the **Quarto notebook** example in the repo's `{r} repo_link("examples folder", "examples/quarto/quarto.qmd")`.

$$
x\in\R
$$

$$
\y=\X\betaB, \beta
$$

# Header

```
format:
  html:
    embed-resources: true
    monofont: Fira Code
    html-math-method: mathjax
    include-in-header:
      - text: <script type="text/javascript" src="./latex.js"></script>
```

[HTML Basics](https://quarto.org/docs/output-formats/html-basics.html)

[embed-resources](https://quarto.org/docs/output-formats/html-basics.html#self-contained)

[html-math-methods](https://quarto.org/docs/output-formats/html-basics.html#latex-equations)

[include-in-header](https://quarto.org/docs/output-formats/html-basics.html#includes)


# Code Chunks and Plots

```{r}
head(mtcars)
```



```{r}
suppressPackageStartupMessages(library(tidyverse))
theme_set(theme_bw())

df <- mtcars
df |>
  ggplot(aes(mpg, disp)) +
  geom_point()
```




# TODO: image preview example

- `{r} image_preview_on_hover("image link text", "img/does_not_exist.PNG")`


# References

- Hadley Wickham books
  - [R for Data Science](https://r4ds.had.co.nz/)
  - [ggplot2](https://ggplot2-book.org/)
  - [Mastering Shiny](https://mastering-shiny.org/)
- More advanced books
  - [R Packages](https://r-pkgs.org/)
  - [Advanced R](https://adv-r.hadley.nz/)
  - [Tidy Modeling with R](https://www.tmwr.org/)
- The Quarto website has notes on [websites](https://quarto.org/docs/websites/), [slides](https://quarto.org/docs/presentations/revealjs/)
- RStudio
  - [Install Fira Code](https://github.com/tonsky/FiraCode/wiki/RStudio-instructions)

