---
title: About the HospitalStarRatings App
format: html
---

This app uses the R package [hospitalstars](https://github.com/jimurick/hospitalstars)

The package is based on two CMS sources:

- Measure score data was taken from the [Hospitals archived data snapshots](https://data.cms.gov/provider-data/archived-data/hospitals)
- The Star rating algorithm was translated to R from the [SAS Package](https://qualitynet.cms.gov/inpatient/public-reporting/overall-ratings/sas)


## Tab 1: Map

I took the data in the package's dataframe `hospitalstars::hospital_info_df` and geocoded it with Google's [geocoding API](https://developers.google.com/maps/documentation/geocoding/overview) in python.


## Tab 2: Plots

CMS collects data from most larger hospitals to calculate the star ratings annually. There are about 50 measures that are used in the calculation; this table gives an example of three:

```{r echo=FALSE, layout="l-body-outset"}
hospitalstars::measure_info_df |>
  dplyr::arrange(desc(report_date)) |>
  dplyr::transmute(
    Group = measure_group, ID = measure_id, Metric = measure_name
  ) |>
  dplyr::slice(c(1, 9, 40)) |>
  knitr::kable(format = "html", table.attr='class="about-table"')
```

The metrics are divided into are 5 groups:

- Mortality
- Patient Experience
- Process
- Readmission
- Safety

## Tab 3: About

Notes to give end-users motivation for the app and information about data sources.

## Gratuitous Math

Standard normal CDF:

$$
\Phi(x)=\int_{-\infty}^{x}\frac{dt}{\sqrt{2\pi e^{t^2}}}
$$

The R function `pnorm()` calculates $\Phi$:

```{r, fig.width=2.5, fig.height=2}
suppressPackageStartupMessages(library(tidyverse))
theme_set(theme_bw())

data.frame(x = seq(from = -3, to = 3, by = 0.01)) |>
  mutate(Φ = pnorm(x)) |>
  ggplot(aes(x, Φ)) + geom_point() + geom_line()
```


